\Large\textbf{3D Graphics Rendering with OpenGL}

### 2019-03-31 Paper.tex
Begin section on VEOs

Graphics rendering is done by defining primitives such as points, lines, and triangles in a 3D space. After these points are defined, their locations are transformed and then they are drawn on the screen through a process called rasterization \cite{mckesson2018}. Figure \ref{fig:triangle-rendering} shows a simple example of the rasterization process. The figure begins by defining a triangle which is composed of three 3D vectors. The rasterizer then determines which pixels the triangle overlaps through a process called scan conversion. One way to do this is to include pixels if their centers are contained within the element. Rasterization then produces fragments, which is an element that contributes to the final color of a pixel \cite{sellers2016}. The fragments are assigned a color in later steps of the rendering process and which might be changed when effects such as lighting are taken into account.

@@

\section{The Rendering Pipeline}
Now that a VAO and VBO have been created, OpenGL is almost ready to draw. The only thing left to do is to tell OpenGL what to do with the data it receives from VAOs, VBOs, and uniforms. When OpenGL receives a draw command, it runs the VAO through a pipeline in order to determine what to draw and how to draw it. This pipeline is responsible for positioning, rasterizing, and coloring primitives. The pipeline is divided into several stages, many of which are configurable through programs written in the OpenGL shading language (GLSL). Figure \ref{fig:pipeline-flow} shows a simplified view of the OpenGL pipeline and Figure \ref{fig:pipeline-visualized} shows how the input is transformed throughout the pipeline. While there are many stages to the pipeline, only two must be defined by the user for rendering to occur. These two stages are the vertex shader and fragment shader. The vertex shader tells OpenGL where a primitive is located, while the fragment shader contributes to the final color of a primitive during rendering.

@@

@@

Once all of this is put together, the application is ready to be compiled and a basic OpenGL program can be executed\footnote{Full program source available at \url{https://github.com/ShepherdJerred-homework/seminar-paper-application}}. This application will display a single triangle with its position and color specified by the code calling OpenGL. The output of the application can be seen in Figure \ref{fig:final-product}.

\section{VEOs}
% TODO VBOs can be repetitive, VEOs allow us to reduce this. Show how to use VEOs with a code sample. Create a simple cube/triangle with VEOs and then render it. Then show the loading of a complex model and render it.

\begin{figure}[h]
    \centering
    \includegraphics[height=6cm]{triangle}
	\caption{A triangle rendered from the vertices in Figure \ref{fig:create-vbo}.}
	\label{fig:final-product}
\end{figure}

>>>>

Graphics rendering is done by defining primitives such as points, lines, and triangles in a 3D space. After these points are defined, their locations are transformed and then they are drawn on the screen through a process called rasterization \cite{mckesson2018}. Figure \ref{fig:triangle-rendering} shows a simple example of the rasterization process. The figure begins by defining a triangle primitive which is composed of three 3D vectors. The rasterizer then determines which fragments the triangle overlaps through a process called scan conversion. One way to do this is to include pixels if their centers are contained within the element. Rasterization then produces list of fragments. A fragment is an element that contributes to the final color of a pixel \cite{sellers2016}. The fragments are assigned a color in later steps of the rendering process and which may be changed when effects such as lighting are taken into account.

@@

% TODO revise
\subsection{VEOs}
Consider the case of drawing a square. This would require two right triangles of equal dimensions connected at their hypotenuses. Each triangle would only have one unique coordinate, with the other two being shared at the hypotenuses, however right now this would have to be sent as six coordinates as show in Figure \ref{fig:square-coords} instead of only four. While these two extra coordinates may seem insignificant, this inefficient use of graphics memory may cause problems, especially when rendering complex scenes with hundreds or thousands of triangles. This problem can be solved with vertex element objects (VEOs).

% TODO go into VEOs and how to use them

\begin{figure}[h]
    \lstinputlisting[language=Java]{code/square.java}
    \caption{The coordinates needed to draw a square, assuming the origin is at the top-left of the screen.}
    \label{fig:square-coords}
\end{figure}

\section{The Rendering Pipeline}
Now that a VAO and VBO have been created, OpenGL is almost ready to draw. The only thing left to do is to tell OpenGL what to do with the data it receives from VAOs, VBOs, and uniforms. When OpenGL receives a draw command, it runs the VAO through a pipeline in order to determine what to draw and how to draw it. This pipeline is responsible for positioning, rasterizing, and coloring primitives. The pipeline is divided into several stages, many of which are configurable through programs written in the OpenGL shading language (GLSL). Figure \ref{fig:pipeline-flow} shows a simplified view of the OpenGL pipeline and Figure \ref{fig:pipeline-visualized} shows how the input is transformed throughout the pipeline. While OpenGL allows several of its stages to be programmed or configured by the user, some of the stages are fixed-function, meaning that OpenGL does not allow any customization to the stage. A couple of notable fixed-function stages include the vertex fetch stage, the primitive assembly stage, and per-sample processing. The vertex fetch stage is the first stage in the pipeline and feeds the pipeline with data from VAOs and VBOs. The primitive assembly stage performs an optimization known as face culling to ensure that only visible triangles are processed. Per-sample processing is the final stage in the OpenGL pipeline and ensures that previously drawn elements are not overwritten unless the new element's depth is higher through a process called depth testing, and writes fragments to the framebuffer to be displayed. % TODO revise

While there are several configurable stages in the pipeline, only two must be defined for rendering to occur. These two stages are the vertex shader and fragment shader. The vertex shader tells OpenGL where a primitive is located, while the fragment shader contributes to the final color of a primitive during rendering.

@@

% TODO go into textures

@@

% TODO revise
Once all of this is put together, the application is ready to be compiled and a basic OpenGL program can be executed\footnote{Full program source available at \url{https://github.com/ShepherdJerred-homework/seminar-paper-application}}. This application will display a single triangle with its position and color specified by the code calling OpenGL. The output of the application can be seen in Figure \ref{fig:final-product}.

% TODO make a more impressive demo
\begin{figure}[h]
    \centering
    \includegraphics[height=6cm]{triangle}
	\caption{A triangle rendered from the vertices in Figure \ref{fig:create-vbo}.}
	\label{fig:final-product}
\end{figure}
